---
title: "ML-Assignment 4"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
date: "2022-11-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(caret)
library(tidyverse)
library(factoextra)
library(esquisse)
set.seed(123)
```


```{r cars}
getwd()
setwd("/Users/thupiliabhinav/Desktop/ML/ML- Assignment 4")
pharma <- read.csv("Pharmaceuticals.csv")

```
#a. Using only the numerical variables (1 to 9) to cluster the 21.
```{r}
norm_mean<- pharma %>% select('Market_Cap', 'Beta', 'PE_Ratio', 'ROE', 'ROA', 'Asset_Turnover', 'Leverage', 'Rev_Growth', 'Net_Profit_Margin')

#Scaling the Data.
norm_train <- preProcess(norm_mean, method = "range")
norm_predict<-predict(norm_train, norm_mean)
```

```{r}
fviz_nbclust(norm_predict, kmeans, method = "wss")
fviz_nbclust(norm_predict, kmeans, method = "silhouette")
```



#From above we calculate kmeans optimal being k=2 
```{r}
k_means_2 <- kmeans(norm_predict, centers = 2, nstart = 25)
k_means_2$centers
```

#Graphical representation of kmeans using cluster:
```{r}
fviz_cluster(k_means_2, data= norm_mean)
```

#Grouping of clusters with original data:
```{r}
k_cluster<- k_means_2$cluster
group_k <- cbind(pharma,k_cluster)
```

#Calculating mean for both clusters:
```{r}
aggregate(group_k[,-c(1,2,12:14)],by=list(group_k$k_cluster),FUN="mean")
```
#b. Interpret the clusters with respect to the numerical variables used in forming the clusters.
# From above we can observe through clustering by "WSS" and "Silhouette" optimal K is 2 
#Cluster 1- has companies with High-Market_Cap, PE_Ratio, ROE, ROA and Net_Profit_Margin.
#Cluster 2- has companies with Low- Market_Cap, PE_Ratio, ROE, ROA and Net_Profit_Margin.

#Grouping of clusters with Original Data:
```{r}
groupk2<-cbind(group_k, pharma$Location,pharma$Exchange)
```

#c.Is there a pattern in the clusters with respect to the numerical variables (10 to 12)?
#With respect to numerical values to columns (10 to 12) are as follows:

#Analysis under column 10.Mediation_recommendation:
#Mediation_recommendation under cluster 1 consists- a) hold recommedations-6 , b) buy recommendations-3, c) sell recommendations - 2
##Mediation_recommendation under cluster 2 consists- a) buy recommendations-5 , b) hold recommedations-3, c) sell recommendations - 2

#Analysis under columns 11.pharma$Location and 12.pharma$Exchange :
#Majority of pharma locations in cluster-1 and cluster-2 are US- based and for pharma exchnage the mojority is NYSE for both the clusters.

#d.Naming for each cluster using any or all of the variables in the dataset:
#Cluster 1- With majority mediation recommendations being hold, this cluster is named as "HOLD CLUSTER".
#Cluster 2- With majority mediation recommendations being buy, this cluster is named as "BUY CLUSTER".




